---
title: "Causes of death in Canadian children 2000-2016"
author: "Tanya Murphy"
date: "20/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://mran.revolutionanalytics.com/snapshot/2018-12-15"))

library(tidyverse)
library(ggpubr)
library(labelled)
library(kableExtra)


```

## Import cause of death data 

Statistics Canada provides an interface to download aggregated data, stratified in many possible ways. These data were downloaded from <https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1310039201&pickMembers%5B0%5D=2.3&pickMembers%5B1%5D=3.1> on Dec 20, 2018 as a CSV file (complete table).

```{r data}

d <- read_csv("13100392-eng/13100392.csv", 
     col_types = cols(REF_DATE = col_integer()))
#summary(d)
#dput(names(d))
names(d) <- c("refdate", "GEO", "DGUID", "agecat", "sex", 
"cause", "measure", "UOM", "UOM_ID", 
"SCALAR_FACTOR", "SCALAR_ID", "VECTOR", "COORDINATE", "value", 
"STATUS", "SYMBOL", "TERMINATED", "DECIMALS")
table(d$agecat, useNA = "ifany")
table(d$Sex, useNA = "ifany")
meta <- read_csv("13100392-eng/cause_codes.csv")
names(meta)

```

## Data preparation

You can also embed plots, for example:

```{r munging}

df <- d %>% 
  filter(
    agecat %in% c(
      "Age at time of death, under 1 year",
      "Age at time of death, 1 to 4 years",
      "Age at time of death, 5 to 9 years",
      "Age at time of death, 10 to 14 years",
      "Age at time of death, 15 to 19 years"
    )
  ) %>% 
  select(refdate, agecat, sex, cause, measure, value) %>% 
  left_join(meta, by = "cause") %>% 
  mutate(agecat = factor(agecat,
                         levels = c(
              "Age at time of death, under 1 year",
      "Age at time of death, 1 to 4 years",
      "Age at time of death, 5 to 9 years",
      "Age at time of death, 10 to 14 years",
      "Age at time of death, 15 to 19 years"             
                         ),
      labels = c("under 1 year", "1 to 4 years", "5 to 9 years",
                       "10 to 14 years", "15 to 19 years"),
      ordered = TRUE),
         sex = factor(sex))

var_label(df$agecat) <- "Age at time of death"
levels(df$sex) <- c("both", "girls", "boys")

a <- df %>% filter(measure == "Number of deaths") %>% 
  group_by(refdate, agecat, sex, mycode) %>% 
  summarise(count = sum(value))

a %>% filter(mycode == "T-All") %>% 
ggplot( aes(y = count, x = sex, col = sex)) +
  geom_bar(stat = "identity") +
  facet_grid(agecat ~ refdate) 


ggplot(a, aes(y = count, x = mycode, col = sex)) +
  geom_bar(stat = "identity")
           

```

## Plots

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.









